Timeliness of security updates for Android

### 2014-08-05 security_scoring.tex
Fix spelling mistake
Updates on reading through

This score is computed from several components:
\begin{description}
  \item[$i$] The proportion of the time were devices exposed to known vulnerabilities. This is equivalent to Acer and Jackson's proposal~\cite{Acer2010} to measure the security based on the proportion of users with at least one unpatched critical vulnerability and is also $1 -$ the Vulnerability Free Days score~\cite{Wright2014}.
  \item[$u$] The proportion of devices run the latest version of Android shipped to any device produced by that manufacturer. This is a measure of internal updatedness, a low score would mean many devices are being left behind.
  \item[$m$] The mean number of outstanding vulnerabilities affecting devices not fixed on any device shipped by the manufacturer. This is related to the Median Active Vulnerabilities measure~\cite{Wright2014} but is the mean rather than the median.
%TODO should we compute the median instead?
\end{description}
\begin{equation}
\mathrm{score} = (1 - i)\times 4 + u \times 3 + \frac{2}{1+e^m} \times 3
\end{equation}

The scores accross the whole of android are that \daMeanInsecurityPerc\ of devices are exposed to known root exploits.
There are on average \daMeanOutstandingVulnerabilities\ outstanding vulnerabilities not fixed on any device.
Only on average \daUpdatednessPerc\ of devices run the most recent version of Android.
This gives a security score of \daSecurityScore/10.
\daTabSecScoressummary
However there are a wide variety of scores depending on the source of the device.
There have been many reports that Google's Nexus devices are better at getting updates than other Android devices because Google makes the original updates and ships them to its devices.
Table~\ref{tab:sec_summary} shows that this is the case with Nexus devices getting much better scores than non-Nexus devices.
\daTabSecScoresmanufacturer
Different manufacturers have very different scores, Table~\ref{tab:sec_manufacturer} shows the scores for the \daNumSigManufacturers\ manufacturers with a significant presence in our data.
Manufacturers are considered significant if we have data from at least \daSigNumDevices\ devices and at least \daSigNumDeviceDays\ days of contributions.
\daTabSecScoresmodel
Even within manufacturers different models can have very different update behaviours and hence security.
Table~\ref{tab:sec_model} shows the results for the \daNumSigModels\ device models which have a significant presence by the same metric.

These tables also show the uncertainty for $i$ and $u$.
$i$ is calculated by taking the total insecure device days and dividing it by the total insecure and secure device days.
The total secure device days and total insecure device days are both counting experiments and so their measurement error is their square root~\cite{Taylor1997}.
Since the numbers involved are large, the uncertainty in $i$ is small.
$u$ is computed by taking the sum of the proportions of devices running the most recent version each day, both the count of devices running the maximum version and total count have sqrt uncertainties.
The Python uncertainties library was used to propagate uncertainties through calculations.

>>>>

This score is computed from several components:
\begin{description}
  \item[$f$] The proportion of the time where devices were free from known critical vulnerabilities. This is equivalent to Acer and Jackson's proposal~\cite{Acer2010} to measure the security based on the proportion of users with at least one unpatched critical vulnerability and is the Vulnerability Free Days score~\cite{Wright2014}.
  \item[$u$] The proportion of devices run the latest version of Android shipped to any device produced by that manufacturer. This is a measure of internal updatedness, a low score would mean many devices are being left behind.
  \item[$m$] The mean number of outstanding vulnerabilities affecting devices not fixed on any device shipped by the manufacturer. This is related to the Median Active Vulnerabilities measure~\cite{Wright2014} but is the mean rather than the median, as this gives a continuous value.
%TODO should we compute the median instead?

We want to provide a score out of 10, since $f$ is the most important metric we weight it more highly.
Since $m$ is an unbounded positive real number, we map it into the range [0--1).
\end{description}
\begin{equation}
\mathrm{score} = f\times 4 + u \times 3 + \frac{2}{1+e^m} \times 3
\end{equation}

The scores across the whole of android are that \daMeanInsecurityPerc\ of devices are exposed to known root exploits.
There are on average \daMeanOutstandingVulnerabilities\ outstanding vulnerabilities not fixed on any device.
Only on average \daUpdatednessPerc\ of devices run the most recent version of Android.
This gives a security score of \daSecurityScore / 10.
\daTabSecScoressummary
However there are a wide variety of scores depending on the source of the device.
There have been many reports that Google's Nexus devices are better at getting updates than other Android devices because Google makes the original updates and ships them to its devices.
Table~\ref{tab:sec_summary} shows that this is the case with Nexus devices getting much better scores than non-Nexus devices.
\daTabSecScoresmanufacturer
Different manufacturers have very different scores, Table~\ref{tab:sec_manufacturer} shows the scores for the \daNumSigManufacturers\ manufacturers with a significant presence in our data with \daSecScoreBestmanufacturer\ (\daSecScoreBestmanufacturerScore\ / 10) doing best and \daSecScoreWorstmanufacturer\ (\daSecScoreWorstmanufacturerScore\ / 10) doing worst.
Manufacturers are considered significant if we have data from at least \daSigNumDevices\ devices and at least \daSigNumDeviceDays\ days of contributions.
\daTabSecScoresmodel
Even within manufacturers different models can have very different update behaviours and hence security.
Table~\ref{tab:sec_model} shows the results for the \daNumSigModels\ device models which have a significant presence by the same metric with \daSecScoreBestmodel\ (\daSecScoreBestmodelScore\ / 10) doing best and \daSecScoreWorstmodel\ (\daSecScoreWorstmodelScore\ / 10) doing worst.

These tables also show the uncertainty for $f$ and $u$.
$f$ is calculated by taking the total secure device days and dividing it by the total insecure and secure device days.
The total secure device days and total insecure device days are both counting experiments and so their measurement error is their square root~\cite{Taylor1997}.
Since the numbers involved are large, the uncertainty in $f$ is small.
$u$ is computed by taking the sum of the proportions of devices running the most recent version each day, both the count of devices running the maximum version and total count have sqrt uncertainties.
The Python \texttt{uncertainties} library was used to propagate uncertainties through calculations.

### 2014-08-05 securityupdates.tex
Initial Acknowledgements
Updates on reading through

\begin{abstract}
Android is the most popular smartphone platform and seeks to provide better security than popular desktop operating systems through increased compartmentalisation and a permission system for apps.
This security relies on the absence of root privilege exploits which allow apps to bypass all protections.
Many such exploits have been found and fixed.
Our hypothesis is that these fixes do not promptly reach the users devices and so much of the time users are running versions of Android known to be vulnerable.
We found that on average over \daMeanInsecurityPerc\ of devices were exposed to known root privilege vulnerabilities and only \daUpdatednessPerc\ of devices run the most recent version of Android.
There was also a period of several months when no devices ran secure versions of Android.
We compare different manufacturers and operators to find which are best at promptly distributing security updates and examine how to improve the situation.
Many phones are sold on long (18 month) contracts with monthly payments yet stop receiving updates before the end of the contract, while Windows XP could be purchased for a one off payment in October 2001 and received security updates until April 2014.
\end{abstract}

@@

Android relies on the Linux kernel to provide compartmentalisation and runs each app as a separate unix `user'.
The Linux kernel is a large piece of software and so there will always be new root exploits found in it~\cite{TODO}.
Manufacturers of Android devices have also added further root exploits when customising it~\cite{Grace2012}.
Between 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a} of malware for Android contained root exploits in 2012.
While this shows that much Android malware does not need a root exploit to work, it can just request the relevant permissions and the user will grant them, a significant proportion does.
This malware is also more dangerous because while ordinary malware can be remotely removed by Google through the Play store, once a root exploit has been used there are no guarantees.

Figure \ref{fig:proportioninsecure} shows the proportion of Android devices in the DeviceAnalyzer data which were running versions of Android known to root exploits at the time.
In \S\ref{sec:background} we will explain where these data comes from and then in \S\ref{sec:results} we will show these data in more detail.

@@

In the past users of Microsoft IIS were advised to switch to a different web server due to the high rate of vulnerability~\cite{Pescatore2001} this pressure helped cause Microsoft to focus on security so that it now does much better~\cite{TODO}.
Part of the reason why deployed devices do not receive updates in a timely manner is that there is little data on what is actually happening and so it is not possible to determine which manufacturers or operators are doing a better job, this data is in~\S\ref{sec:security_scoring}.
Hence there is little incentive for manufacturers and operators to do a better job, we will discuss this further in \S\ref{sec:economics}

@@

\subsection{Root exploits}
We compiled a list of root exploit vulnerabilities for Android containing information on when they were known about, what versions they affected and which versions fixed the problem.
We only looked for \emph{root equivalent} exploits which did not require USB debugging to exploit.
\emph{Root equivalent} means that it grants privileges equivalent in scope to root which can then be used to gain root.
Some phones can be `rooted' by enabling USB debugging and using the special privileges of the adb shell to root the device but only \daAdbPerc\ of devices have USB debugging enabled.
This is not something that applications running on the phone can exploit to break compartmentalisation and so we do not include those exploits.
Some exploits which grant root are not traditional kernel exploits, for example the discovery of flaws in the verification of signatures on Android applications in February 2013~\cite{Forristal2013} meant that applications could pretend to be signed with system keys and hence gain root privileges.

We have published full details of the vulnerabilities listed below in an accompanying technical report~\cite{TODO} and summarise them in Table \ref{tab:andvulns}.
We also have a website\footnote{\url{http://androidvulnerabilities.org/}} which we will keep up to date with future vulnerabilities and which has successfully solicited submissions from others.

There was much manual effort required to track down details of these vulnerabilities as many of them are not in other databases such as the CVE database, the lack of a widely acknowledged unique identifier made identifying whether two reports referenced the same vulnerability difficult.
Previous work has assumed ``any security issue of relevance will eventually get a CVE number assigned"~\cite{Frei2010} which is currently not the case for Android with several root vulnerabilities not having CVE numbers.
For some of the vulnerabilities without CVE numbers we checked with Google whether they had one assigned and they confirmed that there was no number and that they did not intend to get one.

@@

Prior work~\cite{Bilge2012} has shown that after public disclosure exploitation rates increase by 5 orders of magnitude and so from the point of view of widespread danger, the period between public disclosure and the update reaching the user's device is the most critical.
However they also show that zero day exploits are typically used for 312 days before they are publicly disclosed, often to target particular organisations.
Hence when considering vulnerability from the point of view of an organisation which cares about Advanced Persistent Threats, such as those the CESG advice~\cite{CESG2013} is aimed at, the critical period is from when people could have started using it to when the update hits the user device.
So for our vulnerability calculations we use the earliest date we can find recorded evidence that the vulnerability was known about, even if that knowledge might have been confined to a particular manufacturer or hobbyist.
We cannot know if someone reported the vulnerability to the manufacturer and we do know that various agencies are engaged in widespread monitoring of communications and compromise of civilian infrastructure~\cite{TODO} and so they probably know.
Hence we use dates such as when CVEs were registered, when fixes were committed to source repositories, when exploits were committed to repositories as the date first known about.

@@

\subsection{Versions of Android running on devices}
We also need historical information on which versions of Android were running on devices along with other information about those devices which allows us to go into more detail and confirm the reliability of the data our results depend on.
The only suitable data which is available for this is the Device Analyzer data~\cite{Wagner2013}.
The Device Analyzer data contains data from \daNumDevices\ devices with a total of \daDeviceDays\ device days, it contains data for more than \daMonths months for \daMonthsDevices\ devices.
%TODO talk about the Device Analyzer project a bit
Various different kinds of data\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/collected.htm}} are collected by the Device Analyzer Android app\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/}}.
Among these are the build string and API version for the version of Android currently running on the device each day.
The API/SDK version is a well defined integer, however it does not always change with new Android releases, particularly security bug fixes should not increase this value.
The build string is `The user-visible version string', fortunately most (99.9\%) devices in the data have a build string of the form `x.y.z random string' and so it is possible to extract the Android version number.

\subsection{Running a version of Android known to be vulnerable makes a device vulnerable}
While we have not tried to exploit these vulnerabilities on the devices in order to test whether the devices are actually vulnerable, we are confident that most devices.

\section{Results}
\label{sec:results}
In this section we present the results of our analysis showing ????. %TODO summarise our results

@@

This allows us to have confidence in the OS version information.
Figure \ref{fig:os} shows the raw data on the number of devices in the Device Analyzer data running different versions of Android each day.
It shows how old versions are gradually replaced by new ones, the long tail of devices which do not see updates to more recent versions and the fluctuation in the number of devices in the data.
Figure \ref{fig:norm_os} shows the normalised OS version distribution.
\begin{figure}
 \centering
 \includegraphics[width=\columnwidth]{figures/da_os}
 \caption{Android versions in Device Analyzer data over time}
 \label{fig:os}
\end{figure}
\begin{figure}
 \centering
 \includegraphics[width=\columnwidth]{figures/da_norm_os}
 \caption{Android versions in Device Analyzer data over time}
 \label{fig:norm_os}
\end{figure}

@@

\begin{figure}
 \includegraphics[width=0.5\textwidth]{figures/nvulnerabilities_heat.pdf}
 \caption{Proportion of devices affected by different vulnerabilities}
 \label{fig:nvulnerabilities_heat}
\end{figure}
\begin{figure}
 \includegraphics[width=0.5\textwidth]{figures/from_to_updates.pdf}
 \caption{Updates between different Android versions in the DA data}
 \label{fig:from_to_updates}
\end{figure}
\begin{figure}
 \includegraphics[width=0.5\textwidth]{figures/w_security_updates.pdf}
 \caption{Number of updates each week which fixed or may have fixed security vulnerabilities}
 \label{fig:weekly_security_updates}
\end{figure}
Figure \ref{fig:from_to_updates} shows how devices upgrade between different versions of Android.
Device Analyzer (DA) observes upgrade events and the figure shows which versions a device is upgrading from and to.
Mostly the dark cells are above the diagonal and are upgrades.
While many upgrades are from one version to the next version there are also a fair number which skip a few versions.
Surprisingly there are also a small number of downgrade events when older versions of Android are installed on to devices.
Figure \ref{fig:weekly_security_updates} shows the number of devices getting security updates each week, blue shows updates which changed the Android version number from a version with known vulnerabilities to one which had fewer known vulnerabilities.
Red indicates updates which changed the build number but not the version number and so might contain a backported fix for a vulnerability.

\subsection{Update frequency}
%TODO find out what the frequency of updates is

@@

\subsection{Monetisation}
One of the key differences between smartphones and desktops is that it is easier to directly cash out on a smartphone as malware can send premium rate texts or make premium rate phone calls.
It can also steal personal data which already has semantic information associated with it, this also makes it easier to perform ransomware attacks where malware deletes this information from the device and charges a ransom for its return.
It can also do more conventional things such as use the devices for DDoS attacks and altering the ads displayed to send the fees to the malware author.
While many of these things can be done by malware which just requests the relevant permissions, root exploits make this easier, particularly for ransomware and altering the ads displayed.

@@

\section{Evaluation}
Our results rely on the accuracy and representativeness of two data sets, the DeviceAnalyzer data and the AndroidVulnerabilities.org data.

Comparing the API version data from DeviceAnalyzer (Figure~\ref{fig:da_api}) with the ground truth data from Google Play (Figure~\ref{fig:play_api}) using ??? we find that ???.
Since API version is a coarse grained measure of OS version this means that the OS version distribution in DeviceAnalyzer ??? representative.
This shows that our data is ??? useful for examining the behaviour across the whole of Android.
However to compare different manufacturers we need to take small subsets of this data and then compare them.
Since there is no ground truth data available we cannot check definitively

\section{Related work}
\label{sec:related}
We found that since 2011 \daMeanInsecurityPerc\ of devices on average were exposed to known root vulnerabilities, in 2011 Felt et al.\ studied 6 Android handsets and found that they were exposed to root vulnerabilities at least 74\% of the time.
This might indicate that over the last 3 years the security of Android has improved, but the data are not directly comparable.
They also found that 4 of the 46 malware samples (8\%) they analysed contained root exploits, much lower than rates found in later larger studies which found rates of 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a}.
Perhaps the malware authors heard about the results in Felt et al.'s paper?

Finding vulnerabilities is frequently a time consuming manual process but, Brumley et al.~\cite{Brumley2008} showed that it is possible to automatically generate exploits from binary patches.
In a similar way some of the vulnerabilities which have been exploited were publicly found by examining the commit made to the Android which fixed the vulnerability.
This effect has been observed before~\cite{Barth2011} in the Firefox source code repositories, Google does not release the source code until after the release of the update reducing this effect to the delay between an update being released and it reaching the last end user device.
The Woodpecker tool~\cite{Grace2012} automatically finds permission leaks in stock Android phone images.
The update process itself can allow Android apps to gain privilege~\cite{Xing2014} by registering for new permissions before the update which creates that permission is installed.

@@

Various attempts have been made to detect malware.
RiskRanker~\cite{Grace2012a} classified 3,281/118,318 (2.8\%) apps as risky of which 718 (22\%) were malware and 322 (10\%) were previously unknown, an infection rate of 0.6\% across multiple markets. It uses signatures of exploits, static analysis of calls to high risk operations, detection of encrypted native code execution and Dalvik code loading.
DroidRanger~\cite{Zhou2012a} found 148/182,823 (0.08\%) apps to be malicious across multiple markets of which 29 were previously unknown.
It used permission-based behavioural fingerprinting -- looking at the permissions of known malware and heuristic-based filtering -- dynamic loading of both Dalvik and native code.
AnDarwin~\cite{Crussell2013} detects similar apps, it found 169/265,359 (0.06\%) malicious apps based on the difference in permissions between the malicious app and the app which had been cloned.
It used clustering based on semantic vectors derived from the program dependence graphs to detect similar apps.
The Malware Genome project~\cite{Zhou2012b} collected 1,260 malware samples from 2010--2011.
They found that the best case for anti-malware detection on this data was 79.6\%.
However DroidChameleon~\cite{Rastogi2013} found that antivirus products could not detect malware if it was simply automatically permuted.
All these app analysis projects have been hampered by the difficulty of obtaining full datasets of Android apps as Google does not make these available (and researchers who automatically download them, violate Google's Terms of Service).
PlayDrone~\cite{Viennot2014} was a particularly effective project which circumvented Google's protective measures and downloaded over 1,100,000 apps from Google Play, allowing an in depth analysis.

@@

\section{Conclusion}
\label{sec:conclusion}

>>>>

\begin{abstract}
Android is the most popular smartphone platform and seeks to provide better security than popular desktop operating systems through increased compartmentalisation and a permission system for apps.
This security relies on the absence of root privilege exploits which allow apps to bypass protections.
Many such exploits have been found and fixed.
Our hypothesis is that these fixes do not promptly reach the users devices and so much of the time users are running versions of Android known to be vulnerable.
We used Device Analyzer data from over \daOSYearsOfData\ years and \daNumOSDevices\ devices and found that on average over \daMeanInsecurityPerc\ of devices were exposed to known root privilege vulnerabilities and only \daUpdatednessPerc\ of devices run the most recent version of Android.
There was also a period of several months when no devices ran secure versions of Android.
We compare different manufacturers and operators to find which are best at promptly distributing security updates and examine how to improve the situation. %TODO operators analysis
Many phones are sold on long (18 month) contracts with monthly payments yet stop receiving updates before the end of the contract, while Windows XP could be purchased for a one off payment in October 2001 and received security updates until April 2014.
\end{abstract}

@@

Android relies on the Linux kernel to provide compartmentalisation and runs each app as a separate unix `user'.
The Linux kernel is a large piece of software and so there will always be new root exploits found in it~\cite{TODO}.
Manufacturers of Android devices have also added further root exploits when customising it~\cite{Grace2012}.
Between 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a} of malware for Android contained root exploits in 2012.
While this shows that much Android malware does not need a root exploit to work, as it can just request the relevant permissions and the user will grant them, a significant proportion does.
This malware is also more dangerous because while ordinary malware can be remotely removed by Google through the Play store, once a root exploit has been used there are no guarantees.

The proportion of Android devices in the DeviceAnalyzer data which were running versions of Android known to have root vulnerabilities at the time is shown in Figure \ref{fig:proportioninsecure}.
In \S\ref{sec:background} we will explain where these data come from and then in \S\ref{sec:results} we will show these data in more detail.

@@

In the past users of Microsoft IIS were advised to switch to a different web server due to the high rate of vulnerability~\cite{Pescatore2001} this pressure helped cause Microsoft to focus on security so that it now does much better~\cite{TODO}.
Part of the reason why deployed devices do not receive updates in a timely manner is that there is little data on what is actually happening and so it is not possible to determine which manufacturers or operators are doing a better job.
Hence there is little incentive for manufacturers and operators to do a better job, we provide this data  in~\S\ref{sec:security_scoring} and discuss incentives further in \S\ref{sec:economics}

@@

\subsection{Root exploits}
We compiled a list of root exploit vulnerabilities for Android containing information on when they were known about, what versions they affected and which versions fixed the problem.
We only looked for \emph{root equivalent} exploits which did not require USB debugging to exploit.
\emph{Root equivalent} means that it grants privileges equivalent in scope to root which can then be used to gain root.
Some phones can be `rooted' by enabling USB debugging and using the special privileges of the adb shell to root the device but only \daAdbPerc\ of devices have USB debugging enabled.
This is not something that applications running on the phone can exploit to break compartmentalisation and so we do not include those exploits.
Unfortunately many published exploits use adb for convenience and so determining whether that is necessary to exploit the vulnerability can be difficult.
Some exploits which grant root are not traditional kernel exploits, for example the discovery of flaws in the verification of signatures on Android applications in February 2013~\cite{Forristal2013} meant that applications could pretend to be signed with system keys and hence gain root privileges.

We have published full details of the \daNumVulnsUsed\ vulnerabilities used for this analysis in an accompanying technical report~\cite{TODO} and summarise them in Table \ref{tab:andvulns}.
We also have a website\footnote{\url{http://androidvulnerabilities.org/}} which we will keep up to date with future vulnerabilities and which has successfully solicited submissions from others.
In this analysis we only use vulnerabilities which affected all Android devices, not any of the manufacturer or hardware specific vulnerabilities

There was much manual effort required to track down details of these vulnerabilities as many of them are not in other databases such as the CVE database, the lack of a widely acknowledged unique identifier made identifying whether two reports referenced the same vulnerability difficult.
Previous work has assumed ``any security issue of relevance will eventually get a CVE number assigned"~\cite{Frei2010} which is currently not the case for Android with several root vulnerabilities not having CVE numbers.
For some of the vulnerabilities without CVE numbers we checked with Google whether they had one assigned and they confirmed that there was no number and that they did not intend to get one (though they did give us the Android bug number and tell us to use that).

@@

Prior work~\cite{Bilge2012} has shown that after public disclosure exploitation rates increase by 5 orders of magnitude and so from the point of view of widespread danger, the period between public disclosure and the update reaching the user's device is the most critical.
However they also show that zero day exploits are typically used for 312 days before they are publicly disclosed, often to target particular organisations.
Hence when considering vulnerability from the point of view of an organisation which cares about Advanced Persistent Threats, such as those the CESG advice~\cite{CESG2013} is aimed at, the critical period is from when people could have started using the vulnerability to when the update hits the user device.
So for our vulnerability calculations we use the earliest date we can find recorded evidence that the vulnerability was known about, even if that knowledge might have been confined to a particular manufacturer or hobbyist.
We cannot know if someone reported the vulnerability to the manufacturer and we do know that various agencies are engaged in widespread monitoring of communications and compromise of civilian infrastructure~\cite{TODO} and so they probably know.
Hence we use dates such as: when CVEs were registered, when fixes were committed to source repositories, or when exploits were committed to repositories; as the date first known about.

@@

\subsection{Versions of Android running on devices}
We also need historical information on which versions of Android were running on devices along with other information about those devices which allows us to go into more detail and confirm the reliability of the data our results depend on.
The only suitable data which is available for this is the Device Analyzer data~\cite{Wagner2013}.
The Device Analyzer data contains data from \daNumDevices\ devices with a total of \daDeviceDays\ device days, it contains data for more than \daMonths months for \daMonthsDevices\ devices.
%TODO talk about the Device Analyzer project a bit
Various different kinds of data\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/collected.htm}} are collected by the Device Analyzer Android app\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/}}.
Among these are the build string and API version for the version of Android currently running on the device each day.
The API/SDK version is a well defined integer, however it does not always change with new Android releases, particularly security bug fixes should not increase this value.
The build string is `The user-visible version string', fortunately most (99.9\%) devices in the data have a build string of the form `x.y.z random string' and so it is possible to extract the Android version number.
On a large proportion of devices `random string' is a well defined build number but different manufacturers use different schema.

\subsection{Running a version of Android known to be vulnerable makes a device vulnerable}
While we have not tried to exploit these vulnerabilities on the devices in order to test whether the devices are actually vulnerable, we are confident that most devices will be if they are running a vesion known to be vulnerable.

\section{Results}
\label{sec:results}
In this section we present the results of our analysis showing \daMeanInsecurityPerc\ of devices to have root vulnerabilities on average, nexus devices to be better than others, \daSecScoreBestmanufacturer\ to be the best manufacturer and ????. %TODO summarise our results
%TODO comparison of nexus update rates with iOS

@@

This allows us to have confidence in the OS version information.
Figure \ref{fig:norm_os} shows the proportion of devices in the Device Analyzer data running different versions of Android each day.
It shows how old versions are gradually replaced by new ones, the long tail of devices which do not see updates to more recent versions and the fluctuation in the number of devices in the data.
\begin{figure}
 \centering
 \includegraphics[width=\columnwidth]{figures/da_norm_os}
 \caption{Android versions in Device Analyzer data over time}
 \label{fig:norm_os}
\end{figure}

@@

\begin{figure}
 \includegraphics[width=0.5\textwidth]{figures/nvulnerabilities_heat.pdf}
 \caption{Proportion of devices affected by different vulnerabilities}
 \label{fig:nvulnerabilities_heat}
\end{figure}

\subsection{Update frequency}
%TODO find out what the frequency of updates is
\begin{figure}
 \includegraphics[width=0.5\textwidth]{figures/from_to_updates.pdf}
 \caption{Updates between different Android versions in the DA data}
 \label{fig:from_to_updates}
\end{figure}
\begin{figure}
 \includegraphics[width=0.5\textwidth]{figures/w_security_updates.pdf}
 \caption{Number of updates each week which fixed or may have fixed security vulnerabilities}
 \label{fig:weekly_security_updates}
\end{figure}
Figure \ref{fig:from_to_updates} shows how devices upgrade between different versions of Android.
Device Analyzer (DA) observes upgrade events and the figure shows which versions a device is upgrading from and to.
Mostly the dark cells are above the diagonal and are upgrades.
While many upgrades are from one version to the next version there are also a fair number which skip a few versions.
Surprisingly there are also a small number of downgrade events when older versions of Android are installed on to devices.
The number of devices getting security updates each week, is shown in Figure \ref{fig:weekly_security_updates}, blue shows updates which changed the Android version number from a version with known vulnerabilities to one which had fewer known vulnerabilities.
Red indicates updates which changed the build number but not the version number and so might contain a backported fix for a vulnerability.
%TODO summary stats, what does this mean

@@

\subsection{User-Agent data}
TODO

\subsection{Comparison with iOS}
TODO

\subsection{Monetisation}
%TODO redo this section in light of related work
One of the key differences between smartphones and desktops is that it is easier to directly cash out on a smartphone as malware can send premium rate texts or make premium rate phone calls.
It can also steal personal data which already has semantic information associated with it, this also makes it easier to perform ransomware attacks where malware deletes this information from the device and charges a ransom for its return.
It can also do more conventional things such as use the devices for DDoS attacks and altering the ads displayed to send the fees to the malware author.
While many of these things can be done by malware which just requests the relevant permissions, root exploits make this easier, particularly for ransomware and altering the ads displayed.

@@

\section{Related work}
\label{sec:related}
We found that since 2011 \daMeanInsecurityPerc\ of devices on average were exposed to known root vulnerabilities, in 2011 Felt et al.\ studied 6 Android handsets and found that they were exposed to root vulnerabilities at least 74\% of the time.
This might indicate that over the last 3 years the security of Android has not improved, but these data are not directly comparable.
They also found that 4 of the 46 malware samples (8\%) they analysed contained root exploits, much lower than rates found in later larger studies which found rates of 36.7\%~\cite{Zhou2012b} and 40\%~\cite{Zhou2012a}.
Perhaps the malware authors heard about the results in Felt et al.'s paper?
Our results show root exploits are still a severe threat.

Finding vulnerabilities is frequently a time consuming manual process but, Brumley et al.~\cite{Brumley2008} showed that it is possible to automatically generate exploits from binary patches.
In a similar way some of the vulnerabilities which have been exploited were publicly found by examining the commit made to the Android which fixed the vulnerability.
This effect has been observed before~\cite{Barth2011} in the Firefox source code repositories, Google does not release the source code until after the release of the update reducing this effect to the delay between an update being released and it reaching the last end user device.
There have been other attempts to automatically find vulnerabilities.
The Woodpecker tool~\cite{Grace2012} automatically finds permission leaks in stock Android phone images.
The update process itself can allow Android apps to gain privilege through `Pileup' vulnerabilities~\cite{Xing2014} by registering for new permissions before the update which creates that permission is installed.

@@

Various attempts have been made to detect malware.
RiskRanker~\cite{Grace2012a} classified 3,281/118,318 (2.8\%) apps as risky of which 718 (22\%) were malware and 322 (10\%) were previously unknown, an infection rate of 0.6\% across multiple markets.
It uses signatures of exploits, static analysis of calls to high risk operations, detection of encrypted native code execution and dynamic Dalvik code loading to detect risky apps.
DroidRanger~\cite{Zhou2012a} found 148/182,823 (0.08\%) apps to be malicious across multiple markets of which 29 were previously unknown.
It used permission-based behavioural fingerprinting -- looking at the permissions of known malware and heuristic-based filtering -- dynamic loading of both Dalvik and native code.
AnDarwin~\cite{Crussell2013} detects similar apps, it found 169/265,359 (0.06\%) malicious apps based on the difference in permissions between the malicious app and the app which had been cloned.
It used clustering based on semantic vectors derived from the program dependence graphs to detect similar apps.
The Malware Genome project~\cite{Zhou2012b} collected 1,260 malware samples from 2010--2011.
They found that the best case for anti-malware detection on this data was 79.6\%.
However DroidChameleon~\cite{Rastogi2013} found that antivirus products could not detect malware if it was simply automatically permuted.
All these app analysis projects have been hampered by the difficulty of obtaining full datasets of Android apps as Google does not make these available (and researchers who automatically download them, violate Google's Terms of Service).
PlayDrone~\cite{Viennot2014} was a particularly effective project which circumvented Google's protective measures and downloaded over 1,100,000 apps from Google Play, allowing an in depth analysis.

@@

\section{Conclusion}
\label{sec:conclusion}
TODO

\section*{Acknowledgements}
Thanks to David Robertson for helpful advice on statistical analysis.
Thanks to Laurent Simon, Thomas Coudray, Adrian Taylor, Justin Case and Khilan Gudka for reporting vulnerabilities in Android.
